---
title: "R code in Chapter 8"
author: "Andy P. Field"
date: 'Last updated `r format(Sys.Date(), "%d %B %Y")`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{R code in Chapter 8}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)

library(magrittr)
library(dplyr)
library(ggplot2)
```

***

## Usage

This file contains code relevant to chapter 1 of *Discovering Statistics Using R and RStudio* by [Andy Field](https://www.discoveringstatistics.com/). These files contain abridged sections from the book so there are some copyright considerations but I offer them under a [Creative Commons Attribution-NonCommercial-NoDerivatives 4.0 International License](http://creativecommons.org/licenses/by-nc-nd/4.0/), ^[Basically you can use this material for teaching and non-profit activities but do not meddle with it or claim it as your own work.]

***

## Load packages

Remember to load the tidyverse:

```{r, eval = F}
library(tidyverse)
```

## Load the data

Remember to install the package for the book with `library(discovr)`. After which you can load data into a tibble by executing:

`name_of_tib <- discovr::name_of_data`

For example, execute these lines to create the tibbles referred to in the chapter:

```{r}
album_tib <- discovr::album_sales
df_tib <- discovr::df_beta
pub_tib <- discovr::pubs
```

If you want to read the file from the CSV and you have set up your project folder as I suggest in Chapter 1, then the general format of the code you would use is:

```{r, eval = FALSE}
tibble_name <- here::here("../data/name_of_file.csv") %>%
  readr::read_csv() %>%
  dplyr::mutate(
    ...
    code to convert character variables to factors
    ...
  )
```

In which you'd replace `tibble_name` with the name you want to assign to the tibble and change `name_of_file.csv` to the name of the file that you're trying to load. You can use mutate to convert categorical variables to factors. For example, for the `album_sales` data you would load it by executing:

```{r, eval = FALSE}
library(here)

album_tib <- here::here("data/album_sales.csv") %>%
  readr::read_csv()
```


## Self test on df_beta

Load the data and fit the model with all cases.

```{r}
df_tib <- discovr::df_beta

df_lm <- df_tib %>% 
  lm(y ~ x, data = .)

df_lm %>% 
  broom::tidy()
```

Fit the model with case 30 excluded.

```{r}
df_tib %>% 
  dplyr::filter(case != 30) %>% 
  lm(y ~ x, data = .) %>% 
  broom::tidy()
```

Other diagnostics (look at the values for case 30):

```{r}
dfbeta(df_lm)
```



## Jane superbrain influential cases

Load the data and fit the model with all cases.

```{r}
pub_tib <- discovr::pubs

pub_lm <- pub_tib %>% 
  lm(mortality ~ pubs, data = .) 

pub_lm %>% 
  broom::tidy()
```

Save influential cases

```{r}
pub_inf <- pub_lm %>% 
  broom::augment() %>%
  dplyr::rename(
    Residual = .resid,
    `Cook's distance` = .cooksd,
    Leverage = .hat,
  ) %>% dplyr::mutate(
    District = 1:8,
    `DF beta (intercept)` = dfbeta(pub_lm)[, 1] %>% round(., 3),
    `DF beta (pubs)` = dfbeta(pub_lm)[, 2] %>% round(., 3),
  ) %>% 
  dplyr::select(District, Residual, `Cook's distance`, Leverage, `DF beta (intercept)`, `DF beta (pubs)`) %>% 
  round(., 3)

pub_inf
```



## Summary information for album sales
### Summary statistics

```{r}
album_sum <- album_tib %>% 
  tidyr::pivot_longer(
    cols = adverts:image,
    names_to = "Variable",
    values_to = "value"
    ) %>% 
  dplyr::group_by(Variable) %>% 
  dplyr::summarize(
    Mean = mean(value, na.rm = TRUE),
    SD = sd(value, na.rm = TRUE),
    `95% CI Lower` = ggplot2::mean_cl_normal(value)$ymin,
    `95% CI Upper` = ggplot2::mean_cl_normal(value)$ymax,
    Min = min(value, na.rm = TRUE),
    Max = max(value, na.rm = TRUE),
  )
album_sum
```

### Scatterplot of album sales against advertising

```{r}
ggplot2::ggplot(album_tib, aes(adverts, sales)) +
  geom_point(colour = "#5c97bf") +
  coord_cartesian(ylim = c(0, 400), xlim = c(0, 2500)) +
  scale_x_continuous(breaks = seq(0, 2500, 500)) +
  scale_y_continuous(breaks = seq(0, 400, 100)) +
  labs(x = "Advertising budget (thousands)", y = "Album sales (thousands)") +
  geom_smooth(method = "lm", colour = "#d47500", fill = "#d47500", alpha = 0.2) +
  theme_minimal()
```

## Fit a model with one predictor

```{r}
album_lm <- lm(sales ~ adverts, data = album_tib, na.action = na.exclude)
```

get text output for the model:

```{r}
summary(album_lm)
```

Get the confidence intervals for the model parameters old-style:

```{r}
confint(album_lm)
```

Or, better .... use `broom()` to get nicely tabulated fit statistics:

```{r}
broom::glance(album_lm)
```

Finally, you can compute multiple *R*:

```{r}
summary(album_lm)$r.squared %>% 
  sqrt()
```



Also use `broom()` to get nicely tabulated parameter estimates (including confidence intervals):

```{r}
broom::tidy(album_lm, conf.int = TRUE)
```

Impress your friends by rounding values to 3 decimal places using `mutate_if()`

```{r}
broom::tidy(album_lm, conf.int = TRUE) %>%
  dplyr::mutate_if(
    vars(is.numeric(.)),
    list(~round(., 3))
  )
```

Or by using the `pixiedust` package:

```{r}
broom::tidy(album_lm, conf.int = TRUE) %>%
  pixiedust::dust() %>% 
  pixiedust::sprinkle(col = 2:7, round = 3)
```



## Fit a model with several predictor

### Bivariate correlations and scatterplots

Plot the data

```{r}
GGally::ggscatmat(album_tib, columns = c("adverts", "airplay", "image", "sales")) +
  theme_minimal()
```


### Build the model


```{r}
album_full_lm <- lm(sales ~ adverts + airplay + image, data = album_tib, na.action = na.exclude)
```

Or, use the update function:

```{r}
album_full_lm <- update(album_lm, .~. + airplay + image)
```

### Fit statistics

```{r}
broom::glance(album_full_lm)
```

multiple R-square:

```{r}
summary(album_full_lm)$r.squared %>% 
  sqrt()
```

### Compare models

```{r}
anova(album_lm, album_full_lm) %>% 
  broom::tidy()
```

### Parameter estimates

As text:

```{r}
summary(album_full_lm)
confint(album_full_lm)
```

As a nicely formatted table:

```{r}
broom::tidy(album_full_lm, conf.int = TRUE)
```

Restrict the output to 3 decimal places

```{r}
broom::tidy(album_full_lm, conf.int = TRUE) %>% 
  dplyr::mutate_if(
    vars(is.numeric(.)),
    list(~round(., 3))
  )
```

### Standardized betas


```{r}
parameters::model_parameters(album_full_lm, standardize = "refit", digits = 3)
```


### Influence measures

```{r}
album_full_rsd <- album_full_lm %>% 
  broom::augment() %>% 
  tibble::rowid_to_column(var = "case_no")

album_full_rsd
```




```{r}
album_full_inf <- influence.measures(album_full_lm)

album_full_inf <- album_full_inf$infmat %>% 
  tibble::as_tibble() %>%
  tibble::rowid_to_column(var = "case_no")
  

album_full_inf
```

If you want to keep all of your diagnostic stats in one object:

```{r}
album_full_rsd <- album_full_rsd %>% 
  dplyr::left_join(., album_full_inf, by = "case_no") %>% 
  dplyr::select(-c(cook.d, hat))
```

### Variance inflation factor

```{r}
car::vif(album_full_lm)
1/car::vif(album_full_lm)
car::vif(album_full_lm) %>% 
  mean()
```

### Casewise diagnostics

```{r}
get_cum_percent <- function(var,  cut_off = 1.96){
  ecdf_var <- abs(var) %>% ecdf()
  100*(1 - ecdf_var(cut_off))
}


album_full_rsd %>%
  dplyr::summarize(
    `z >= 1.96` = get_cum_percent(.std.resid),
    `z >= 2.58` = get_cum_percent(.std.resid, cut_off = 2.58),
    `z >= 3.29` = get_cum_percent(.std.resid, cut_off = 3.29)
  )
```

```{r}
album_full_rsd %>% 
  dplyr::filter(abs(.std.resid) >= 1.96) %>%
  dplyr::select(case_no, .std.resid, .resid) %>%
  dplyr::arrange(.std.resid)
```

```{r}
album_full_inf %>%
  dplyr::arrange(desc(cook.d)) %>% 
  dplyr::mutate_if(
    vars(is.numeric(.)),
    list(~round(., 3))
  )
```

Show case with any df beta with an absolute value greater than 1.

```{r}
album_full_inf %>% 
  dplyr::filter_at(
    vars(starts_with("dfb")),
    any_vars(abs(.) > 1)
  ) %>% 
  dplyr::select(case_no, starts_with("dfb")) %>% 
  dplyr::mutate_if(
    vars(is.numeric(.)),
    list(~round(., 3))
  )
```

Show cases with problematic leverage or CVR:

```{r}
leverage_thresh <- 3*mean(album_full_inf$hat, na.rm = TRUE)                

album_full_inf %>%
  dplyr::filter(
    hat > leverage_thresh | !between(cov.r, 1 - leverage_thresh, 1 + leverage_thresh)
    ) %>%  
  dplyr::select(case_no, cov.r, hat, cook.d) %>% 
  dplyr::mutate_if(
    vars(is.numeric(.)),
    list(~round(., 3))
  )

  
```

### Diagnostic plots

Standard influence plots

```{r}
plot(album_full_lm, which = 4:6)
```

Residual plot:

```{r}
plot(album_full_lm, which = c(1, 3))
```

Q-Q plot:

```{r}
plot(album_full_lm, which = 2)
```

### Prettier diagnostic plots

The `ggfortify` package allows us to produce prettier versions of the diagnostic plots. For example:

 influence plots

```{r}
library(ggfortify)
ggplot2::autoplot(album_full_lm, which = 4:6,
                  colour = "#5c97bf",
                  smooth.colour = "#ef4836",
                  alpha = 0.5,
                  size = 1) + 
  theme_minimal()
```

Residual plot:

```{r}
ggplot2::autoplot(album_full_lm, which = c(1, 3),
                  colour = "#5c97bf",
                  smooth.colour = "#ef4836",
                  alpha = 0.5,
                  size = 1) + 
  theme_minimal()
```

Q-Q plot:

```{r}
ggplot2::autoplot(album_full_lm, which = 2,
                  colour = "#5c97bf",
                  smooth.colour = "#ef4836",
                  alpha = 0.5,
                  size = 1) + 
  theme_minimal()

```



### Robust linear model

Robust estimates:

```{r}
album_full_rob <- robust::lmRob(sales ~ adverts + airplay + image, data = album_tib, na.action = na.exclude)
summary(album_full_rob)
```

Robust standard errors and confidence intervals:

```{r}
parameters::model_parameters(
  album_full_lm,
  robust = TRUE,
  vcov.type = "HC4",
  digits = 3
  )
```

Bootstrap

```{r}
parameters::model_parameters(
  album_full_lm,
  bootstrap = TRUE,
  digits = 3
  )
```

### Bayes factors

```{r}
album_bf <- album_tib %>%
  BayesFactor::regressionBF(sales ~ adverts + airplay + image, rscaleCont = "medium", data = .)

album_bf
```

### Bayesian parameter estimates

```{r}
album_full_bf <- album_tib %>%
  BayesFactor::lmBF(sales ~ adverts + airplay + image, rscaleCont = "medium", data = .)

album_full_post <- BayesFactor::posterior(album_full_bf, iterations = 10000)
summary(album_full_post)
```

