---
title: "R code in Chapter 2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{R code in Chapter 2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

```{r setup, results="hide"}
# Make sure to load these packages

library(discovr)
library(tidyverse)
```

## Usage

This file contains code relevant to chapter 1 of *Discovering Statistics Using R and RStudio* by [Andy Field](https://www.discoveringstatistics.com/). These files contain abridged sections from the book so there are some copyright considerations but I offer them under a [Creative Commons Attribution-NonCommercial-NoDerivatives 4.0 International License](http://creativecommons.org/licenses/by-nc-nd/4.0/), ^[Basically you can use this material for teaching and non-profit activities but do not meddle with it or claim it as your own work.]

## The median

```{r median}

fb_tib <- tibble::tibble(.rows = 11) %>% 
  dplyr::mutate(
    friends = c(57, 40, 103, 234, 93, 53, 116, 98, 108, 121, 22)
  )


# Base R:

median(fb_tib$friends)

# One pipe:

fb_tib %>%
  dplyr::summarize(
    median =  median(friends)
  )

```

## The mean

```{r mean}
# Base R:

mean(fb_tib$friends)
mean(fb_tib$friends, trim = 0.1)

# A pipe to get the mean, trimmed mean and median:

fb_tib %>%
  dplyr::summarize(
    median =  median(friends),
    mean =  mean(friends),
    `trimmed mean 10%` =  mean(friends, trim = 0.1)
  )
```

###	The dispersion in a distribution 

```{r dispersion}

# The range
max(fb_tib$friends) - min(fb_tib$friends)

# Quartiles
quantile(fb_tib$friends, probs = c(0.25, 0.5, 0.75))

# Lower quartile
quantile(fb_tib$friends, probs = 0.25)

# Upper quartile
quantile(fb_tib$friends, probs = 0.75)

# Inter-quartile range
IQR(fb_tib$friends)

# Variance
var(fb_tib$friends)

# standard deviation
sd(fb_tib$friends)

# Tidyverse sumptuousness:

fb_tib %>%
  dplyr::summarize(
    median =  median(friends),
    mean =  mean(friends),
    `trimmed mean 10%` =  mean(friends, trim = 0.1),
    range = max(friends) - min(friends),
    `lower quartile` = quantile(friends, probs = 0.25),
    `upper quartile` = quantile(friends, probs = 0.75),
    IQR = IQR(friends),
    var = var(friends),
    sd = sd(friends)
)

```


***

## Pieces of great

### Pieces of great 2.1

```{r}
round(3.211420)
round(3.211420, 2)
round(3.211420, 4)

round(mean(fb_tib$friends, trim = 0.1), 2)
mean(fb_tib$friends, trim = 0.1) %>% round(., 2)


fb_tib %>%
  dplyr::summarize(
    median =  median(friends),
    mean =  mean(friends),
    `trimmed mean 10%` =  mean(friends, trim = 0.1)
  ) %>% 
  round(., 2)

```

