---
title: "R code in Chapter 9"
author: "Andy P. Field"
date: 'Last updated `r format(Sys.Date(), "%d %B %Y")`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{R code in Chapter 9}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)

library(magrittr)
library(dplyr)
library(ggplot2)
```



***

## Usage

This file contains code relevant to chapter 1 of *Discovering Statistics Using R and RStudio* by [Andy Field](https://www.discoveringstatistics.com/). These files contain abridged sections from the book so there are some copyright considerations but I offer them under a [Creative Commons Attribution-NonCommercial-NoDerivatives 4.0 International License](http://creativecommons.org/licenses/by-nc-nd/4.0/), ^[Basically you can use this material for teaching and non-profit activities but do not meddle with it or claim it as your own work.]

***

## Load packages

Remember to load the tidyverse:

```{r, eval = F}
library(tidyverse)
```

## Load the data

Remember to install the package for the book with `library(discovr)`. After which you can load data into a tibble by executing:

`name_of_tib <- discovr::name_of_data`

For example, execute these lines to create the tibbles referred to in the chapter:

```{r}
cloak_tib <- discovr::invisibility_cloak
cloak_rm_tib <- discovr::invisibility_rm
```

If you want to read the file from the CSV and you have set up your project folder as I suggest in Chapter 1, then the general format of the code you would use is:

```{r, eval = FALSE}
tibble_name <- here::here("../data/name_of_file.csv") %>%
  readr::read_csv() %>%
  dplyr::mutate(
    ...
    code to convert character variables to factors
    ...
  )
```

In which you'd replace `tibble_name` with the name you want to assign to the tibble and change `name_of_file.csv` to the name of the file that you're trying to load. You can use mutate to convert categorical variables to factors. For example, for the `invisibility_cloak` data you would load it by executing:

```{r, eval = FALSE}
library(here)

cloak_tib <- here::here("data/invisibility.csv") %>%
  readr::read_csv() %>%
  dplyr::mutate(
    cloak = forcats::as_factor(cloak)
  )
```

## Entering data

You were given the task to enter the data, there are several ways to do this, but here's one of them:

```{r}
cloak_tib <- tibble::tibble(
    id = c("Kiara", "Anupama", "Collin", "Ian", "Vanessa", "Darrell", "Tyler", "Steven", "Katheryn", "Kirkpatrick", "Melissa", "Kinaana", "Shajee'a", "Sage", "Jorge", "Conan","Tamara","Man", "Joseph", "Israa", "Kathryn", "Luis", "Devante", "Jerry"),
    cloak = gl(2, 12, labels = c("No cloak", "Cloak")),
    mischief = c(3, 1, 5, 4, 6, 4, 6, 2, 0, 5, 4, 5, 4, 3, 6, 6, 8, 5, 5, 4, 2, 5, 7, 5)
  )

```

## Descriptive statistics

```{r}
cloak_tib %>% 
  dplyr::group_by(cloak) %>% 
  dplyr::summarize(
    n = n(),
    mean = mean(mischief),
    sd = sd(mischief),
    ci_lower = ggplot2::mean_cl_normal(mischief)$ymin,
    ci_upper = ggplot2::mean_cl_normal(mischief)$ymax,
    iqr = IQR(mischief),
    skew = moments::skewness(mischief),
    kurtosis = moments::kurtosis(mischief)
  )
```

```{r}
cloak_lm <- lm(mischief ~ cloak, data = cloak_tib)
summary(cloak_lm)
```

## t-test from summary data

```{r}

t_from_means <- function(m1, m2, sd1, sd2, n1, n2){
  df <- n1 + n2 - 2
	poolvar <- (((n1-1)*sd1^2)+((n2-1)*sd2^2))/df
	t <- (m1-m2)/sqrt(poolvar*((1/n1)+(1/n2)))
	sig <- 2*(1-(pt(abs(t),df)))
	paste0("t(df = ", df, ") = ", round(t, 3), ", p = ", round(sig, 5))
}

t_from_means(m1 = 5, m2 = 3.75, sd1 = 1.651, sd2 = 1.913, n1 = 12, n2 = 12)

```

## Error bar plot

```{r}
ggplot2::ggplot(cloak_tib, aes(cloak, mischief)) +
  geom_violin() +
  stat_summary(fun.data = "mean_cl_normal") +
  labs(x = "Cloak group", y = "Acts of mischief") +
  theme_minimal()

```

## Independent t-test
```{r}
cloak_mod <- t.test(mischief ~ cloak, data = cloak_tib)
cloak_mod 
```

## Robust tests

### `yuen()`

```{r}
WRS2::yuen(mischief ~ cloak, data = cloak_tib)
WRS2::yuen(mischief ~ cloak, data = cloak_tib, tr = 0.1)
```

### `yuenbt()`
```{r}
WRS2::yuenbt(mischief ~ cloak, data = cloak_tib, nboot = 1000, side = TRUE)
```
### `pb2gen()`

```{r}
WRS2::pb2gen(mischief ~ cloak, data = cloak_tib, nboot = 1000)
```

## Bayes factor for indepedent means

```{r}
cloak_bf <- BayesFactor::ttestBF(formula = mischief ~ cloak, data = cloak_tib)
cloak_bf
```

```{r}
BayesFactor::posterior(cloak_bf, iterations = 1000) %>% 
  summary()
```

```{r}
1/cloak_bf
```

## Effect sizes

```{r}
effectsize::t_to_r(t = cloak_mod$statistic, df_error = cloak_mod$parameter)

effectsize::cohens_d(mischief ~ cloak, data = cloak_tib)
effectsize::hedges_g(mischief ~ cloak, data = cloak_tib)
effectsize::glass_delta(mischief ~ cloak, data = cloak_tib)
```

```{r}

r_from_t <- function(t, df, digits = 5){
  r <- sqrt(t^2/(t^2+df))
  r <- round(r, digits)
  return(paste0("r = ", r))
}

r_from_t(t = cloak_mod$statistic, df = cloak_mod$parameter)

r_from_t(t = cloak_mod$statistic, df = cloak_mod$parameter, digits = 2)
```

## Entering data for related means

```{r}
cloak_rm_tib <- tibble::tibble(
    id = rep(c("Alia", " Elisia", "Dominika", "Noah", "Haashid", "Zainab", "Reyna", "Victoria", "Najaat", "Jordyn", "Zaki", "Rajohn"), 2),
    cloak = gl(2, 12, labels = c("No cloak", "Cloak")),
    mischief = c(3, 1, 5, 4, 6, 4, 6, 2, 0, 5, 4, 5, 4, 3, 6, 6, 8, 5, 5, 4, 2, 5, 7, 5)
  )
```

Or load it from the package:

```{r}
cloak_rm_tib <- discovr::invisibility_rm
```

Plot of the differences between scores:

```{r}
cloak_rm_tib %>% 
  tidyr::pivot_wider(
    id_cols = id,
    values_from = mischief,
    names_from = cloak
    ) %>% 
  dplyr::mutate(
    diff = Cloak - `No cloak`
  ) %>%  ggplot2::ggplot(., aes(sample = diff)) +
    qqplotr::stat_qq_band(fill = "#5c97bf", alpha = 0.3) +
    qqplotr::stat_qq_line(colour = "#5c97bf") +
    qqplotr::stat_qq_point() +
    labs(x = "Theoretical quantiles", y = "Sample quantiles") +
    theme_minimal()
```

## Paired *t*-test

```{r}
cloak_rm_mod <- t.test(mischief ~ cloak, data = cloak_rm_tib, paired = TRUE)
cloak_rm_mod 
```

Showing that order matters:

```{r}
cloak_rm_tib %>% 
  dplyr::sample_n(24) %>%
  t.test(mischief ~ cloak, data = ., paired = TRUE)
```
 
Making sure that data are in the correct order by sorting by participant id:

```{r}
cloak_rm_tib %>% 
  dplyr::arrange(id) %>%
  t.test(mischief ~ cloak, data = ., paired = TRUE)
```

## Robust tests of dependent means
### `yuend()`

```{r}
cloak_messy_tib <- cloak_rm_tib %>% 
  tidyr::spread(value = mischief, key = cloak)

WRS2::yuend(cloak_messy_tib$Cloak, cloak_messy_tib$`No cloak`)
```

## Bayes factor for depedent means

```{r}
cloak_rm_bf <- BayesFactor::ttestBF(cloak_messy_tib$Cloak, cloak_messy_tib$`No cloak`, paired = TRUE, rscale = "medium")
cloak_rm_bf
```

```{r}
BayesFactor::posterior(cloak_rm_bf, iterations = 1000) %>% 
  summary()
```

```{r}
1/cloak_rm_bf
```

## Effect sizes for dependent means

Do the same as for independent designs (see earlier). If you must, you can do this:

```{r}
cloak_rm_tib %>% 
  dplyr::arrange(id) %>% 
  effectsize::cohens_d(mischief ~ cloak, data = ., paired = TRUE)

cloak_rm_tib %>% 
  dplyr::arrange(id) %>%
  effectsize::hedges_g(mischief ~ cloak, data = ., paired = TRUE)


effectsize::t_to_r(t = cloak_mod$statistic, df_error = cloak_mod$parameter)
```
