---
title: "R code in Chapter 10"
author: "Andy P. Field"
date: 'Last updated `r format(Sys.Date(), "%d %B %Y")`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{R code in Chapter 10}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```


***

## Usage

This file contains code relevant to chapter 1 of *Discovering Statistics Using R and RStudio* by [Andy Field](https://www.discoveringstatistics.com/). These files contain abridged sections from the book so there are some copyright considerations but I offer them under a [Creative Commons Attribution-NonCommercial-NoDerivatives 4.0 International License](http://creativecommons.org/licenses/by-nc-nd/4.0/), ^[Basically you can use this material for teaching and non-profit activities but do not meddle with it or claim it as your own work.]

***

## Load packages

Remember to load the relevant packages mentioned at the start of the chapter:

```{r}

library(tidyverse)

```

## Load the data

Remember to load the package for the book with `library(discovr)`. After which you can load data into a tibble by executing:

`name_of_tib <- discovr::name_of_data`

For example, execute these lines to create the tibbles referred to in the chapter:

```{r}
vids_tib <- discovr::video_games
infidelity_tib <- discovr::lambert_2012
newz_tib <- discovr::bronstein_2019
newz_md_tib <- discovr::bronstein_miss_2019
```

If you want to read the file from the CSV and you have set up your project folder as I suggest in Chapter 1, then the general format of the code you would use is:

```{r, eval = FALSE}
tibble_name <- here::here("../data/name_of_file.csv") %>%
  readr::read_csv() %>%
  dplyr::mutate(
    ...
    code to convert character variables to factors
    ...
  )
```

In which you'd replace `tibble_name` with the name you want to assign to the tibble and change `name_of_file.csv` to the name of the file that you're trying to load. You can use mutate to convert categorical variables to factors. For example, for the `video_games` data you would load it by executing:

```{r, eval = FALSE}
library(here)

vids_tib <- here::here("data/video_games.csv") %>%
  readr::read_csv()
```

## Centering variables

A basic method

```{r eval = FALSE}

vids_tib <- discovr::video_games

vids_tib <- vids_tib %>% 
  dplyr::mutate(
    caunts_cent = caunts - mean(caunts, na.rm = TRUE),
    vid_game_cent = vid_game - mean(vid_game, na.rm = TRUE)
  )

vids_tib
```


```{r}
# Create a general function to do the centring
centre <- function(var){
  var - mean(var, na.rm = TRUE)
}

# use the general function to centre multiple variables at once
vids_tib <- vids_tib %>% 
  dplyr::mutate_at(
    vars(vid_game, caunts),
    list(cent = centre)
  )

vids_tib
```

```{r, eval = FALSE}

vids_tib <- vids_tib %>% 
  dplyr::mutate(
    interaction = caunts_cent*vid_game_cent
  )

vids_tib
```

## Fitting the model

```{r}
aggress_lm <- lm(aggress ~ caunts_cent + vid_game_cent + caunts_cent:vid_game_cent, data = vids_tib)
```

Or

```{r eval = FALSE}
aggress_lm <- lm(aggress ~ caunts_cent*vid_game_cent, data = vids_tib)
```

Viewing the model:

```{r}
broom::glance(aggress_lm)
broom::tidy(aggress_lm, conf.int = TRUE)
```

Or, also round the digits 

```{r}
aggress_lm %>%
  broom::tidy() %>% 
  dplyr::mutate_if(
    vars(is.numeric(.)),
    list(~round(., digits = 3))
  )
```

Fit a robust model:

```{r}
parameters::model_parameters(aggress_lm, robust = TRUE, vcov.type = "HC4", digits = 3)
```


## Simple slopes and Johnson-Neyman interval

```{r}
interactions::sim_slopes(
  aggress_lm,
  pred = vid_game_cent,
  modx = caunts_cent,
  jnplot = TRUE,
  robust = TRUE,
  confint = TRUE
  )

```

Simple slopes plot

```{r}
interactions::interact_plot(
  aggress_lm,
  pred = vid_game_cent,
  modx = caunts_cent,
  interval = TRUE,
  x.label = "Time playing video games per week (hours)",
  y.label = "Predicted agression",
  legend.main = "Callous unemotional traits"
  ) 
```

# Mediation
## Self-test

```{r}
m1 <- lm(phys_inf ~ ln_porn, data = infidelity_tib)
m2 <- lm(commit ~ ln_porn, data = infidelity_tib)
m3 <- lm(phys_inf ~ ln_porn + commit, data = infidelity_tib)

broom::tidy(m1, conf.int = TRUE)
broom::tidy(m2, conf.int = TRUE)
broom::tidy(m3, conf.int = TRUE) %>% mutate_if(vars(is.numeric(.)), list(~round(., 3)))


```

## The better method

Look at patterns of missing data

```{r}

infidelity_tib %>% 
  dplyr::select(id, commit, ln_porn, phys_inf) %>% 
  mice::md.pattern()

infidelity_tib %>% 
  dplyr::select(id, commit, ln_porn, phys_inf) %>%
  mice::ic()


infidelity_tib <- infidelity_tib %>% 
  dplyr::select(id, commit, ln_porn, phys_inf) %>% 
  na.omit()

```

Fit the model

```{r}

infidelity_mod <- 'phys_inf ~ c*ln_porn + b*commit
                   commit ~ a*ln_porn

                   indirect_effect := a*b
                   total_effect := c + (a*b)
                   '

infidelity_fit <- lavaan::sem(infidelity_mod, data = infidelity_tib, missing = "FIML", estimator = "MLR")

broom::glance(infidelity_fit)

broom::tidy(infidelity_fit, conf.int = TRUE) %>%
  dplyr::mutate_if(
    vars(is.numeric(.)),
    list(~round(., 3))
  )
```

Non-robust version for shits and giggles:

```{r}
infidelity_nonrob <- lavaan::sem(infidelity_mod, data = infidelity_tib, missing = "FIML")

broom::tidy(infidelity_rob, conf.int = TRUE) %>%
  dplyr::mutate_if(
    vars(is.numeric(.)),
    list(~round(., 3))
  )
```

Standardized solution:

```{r}
lavaan::standardizedsolution(infidelity_fit)
```


## Two mediators

```{r}
newz_mod <- 'fake_newz ~ c*delusionz + b1*thinkz_open + b2*thinkz_anal
                thinkz_open ~ a1*delusionz
                thinkz_anal ~ a2*delusionz

                indirect_open := a1*b1
                indirect_anal := a2*b2
                total_effect := c + (a1*b1) + (a2*b2)
                thinkz_open ~~ thinkz_anal
                '

newz_fit <- lavaan::sem(newz_mod, data = newz_tib, se = "bootstrap")

 broom::tidy(newz_fit, conf.int = TRUE) %>%
  dplyr::mutate_if(
    vars(is.numeric(.)),
    list(~round(., 3))
  )
```

# Missing data

## Look for patterns of missingness

```{r}
newz_md_tib %>%
  dplyr::select(-id) %>% 
  mice::md.pattern(., rotate.names = TRUE)
```

```{r}
newz_md_tib %>%
  mice::ic()
```


## Casewise deletion

```{r}

fake_lm <- lm(fake_newz ~ delusionz + thinkz_open + thinkz_anal, data = newz_md_tib)

broom::glance(fake_lm)

broom::tidy(fake_lm, conf.int = TRUE) %>% 
  dplyr::mutate_if(
    vars(is.numeric(.)),
    list(~round(., 3))
    )
```






complete data

```{r}
lm(fake_newz ~ delusionz + thinkz_open + thinkz_anal, data = newz_tib) %>%
  broom::tidy() %>% 
    dplyr::mutate_if(
    vars(is.numeric(.)),
    list(~round(., 3))
  )
```

## FIML

```{r}
newz_lm <- 'fake_newz ~ delusionz + thinkz_open + thinkz_anal
            fake_newz ~ 1
            '

# don't need the intercept because variables are standardized, otherwise include fake_newz ~ 1)

newz_lm_fit <- lavaan::sem(newz_lm, data = newz_tib)

 broom::tidy(newz_lm_fit, conf.int = TRUE) %>%
  dplyr::mutate_if(
    vars(is.numeric(.)),
    list(~round(., 3))
  )
 
newz_lm <- 'fake_newz ~ delusionz + thinkz_open + thinkz_anal
            '
newz_lm_fit <- lavaan::sem(newz_lm, data = newz_md_tib, missing = "FIML")

 broom::tidy(newz_lm_fit, conf.int = TRUE) %>%
  dplyr::mutate_if(
    vars(is.numeric(.)),
    list(~round(., 3))
  ) 
 


```

## Multiple imputation



Imputation

```{r}
newz_imps <- newz_md_tib %>% 
  dplyr::select(-id) %>% 
  mice::mice(method = "pmm", m = 20)



newz_mi_fit <- with(newz_imps, lm(fake_newz ~ delusionz + thinkz_open + thinkz_anal))

newz_pool_fit <- mice::pool(newz_mi_fit)
summary(newz_pool_fit)
```

