---
title: "R code in Chapter 11"
author: "Andy P. Field"
date: 'Last updated `r format(Sys.Date(), "%d %B %Y")`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{R code in Chapter 11}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

```{r setup}

```

## Usage

This file contains code relevant to chapter 1 of *Discovering Statistics Using R and RStudio* by [Andy Field](https://www.discoveringstatistics.com/). These files contain abridged sections from the book so there are some copyright considerations but I offer them under a [Creative Commons Attribution-NonCommercial-NoDerivatives 4.0 International License](http://creativecommons.org/licenses/by-nc-nd/4.0/), ^[Basically you can use this material for teaching and non-profit activities but do not meddle with it or claim it as your own work.]



```{r}


puppies_dummy <- puppies %>%
  dplyr::mutate(
    dummy1 = ifelse(dose == "30 mins", 1, 0),
    dummy2 = ifelse(dose == "15 mins", 1, 0)
  )

puppies_contrast <- puppies %>%
  dplyr::mutate(
    dummy1 = ifelse(dose == "Control", -2/3, 1/3),
    dummy2 = ifelse(dose == "Control", 0, 
                    ifelse(dose == "15 mins", -0.5, 0.5))
  )
```

