---
title: "R code in Chapter 5"
author: "Andy P. Field"
date: 'Last updated `r format(Sys.Date(), "%d %B %Y")`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{R code in Chapter 5}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

```{r setup}
library(tidyverse)
library(GGally)
```

## Usage

This file contains code relevant to chapter 5 of *Discovering Statistics Using R and RStudio* by [Andy Field](https://www.discoveringstatistics.com/). These files contain abridged sections from the book so there are some copyright considerations but I offer them under a [Creative Commons Attribution-NonCommercial-NoDerivatives 4.0 International License](http://creativecommons.org/licenses/by-nc-nd/4.0/), ^[Basically you can use this material for teaching and non-profit activities but do not meddle with it or claim it as your own work.]

## Load the data

Remember to load the package with `library(discovr)`. After which you can load data into a tibble by executing:

`name_of_tib <- discovr::name_of_data`

For example, execute these lines to create the three tibbles referred to in the chapter:

```{r}
cat_tib <- discovr::catterplot
exam_tib <- discovr::exam_anxiety
grammar_tib <- discovr::grammar
notebook_tib <- discovr::notebook
hiccups_tib <- discovr::hiccups
ong_tib <- discovr::ong_tidy
wish_tib <- discovr::jiminy_cricket
```

If you want to read the file from the CSV and you have set up your project folder as I suggest in Chapter 1, then the general format of the code you would use is:

```{r, eval = FALSE}
tibble_name <- here::here("../data/name_of_file.csv") %>%
  readr::read_csv() %>%
  dplyr::mutate(
    ...
    code to convert character variables to factors
    ...
  )
```

In which you'd replace `tibble_name` with the name you want to assign to the tibble and change `name_of_file.csv` to the name of the file that you're trying to load. You can use mutate to convert categorical variables to factors. For example, for the `notebook` data you would load it by executing:

```{r, eval = FALSE}
library(here)

notebook_tib <- here::here("data/notebook.csv") %>%
  readr::read_csv() %>%
  dplyr::mutate(
    sex = forcats::as_factor(sex),
    film = forcats::as_factor(film)
  )
```


## Histograms

```{r}
ggplot2::ggplot(wish_tib, aes(success)) +
 geom_histogram(binwidth = 5, fill = "#56B4E9", colour = "#336c8b", alpha = 0.2) +
 labs(y = "Frequency", x = "Success scores (0-100)", title = "Histogram of success scores") +
 theme_minimal()
```


```{r}
ggplot2::ggplot(wish_tib, aes(success)) +
 geom_histogram(binwidth = 5, fill = "#56B4E9", colour = "#336c8b", alpha = 0.2) +
 labs(y = "Frequency", x = "Success scores (0-100)", title = "Histogram of success scores") +
 facet_grid(strategy~time) +
 theme_minimal()
```

## Frequency polygon

```{r}
ggplot2::ggplot(wish_tib, aes(success)) +
 geom_freqpoly(binwidth = 5, colour = "#336c8b") +
 labs(y = "Frequency", x = "Success scores (0-100)", title = "Histogram of success scores") +
 facet_grid(strategy~time) +
 theme_minimal()
```

## Boxplots

```{r}
ggplot2::ggplot(wish_tib, aes(time, success)) +
  geom_boxplot(fill = "#5c97bf", alpha = 0.7) +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
  labs(x = "Time", y = "Success (out of 100)") +
  theme_minimal()
```

```{r}
ggplot2::ggplot(wish_tib, aes(time, success)) +
  geom_boxplot(fill = "#5c97bf", alpha = 0.7) +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
  labs(x = "Time", y = "Success (out of 100)") +
  facet_wrap(~ strategy) +
  theme_minimal()
```

```{r}
ggplot2::ggplot(wish_tib, aes(time, success, fill = strategy)) +
  geom_boxplot(alpha = 0.7) +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
  labs(x = "Time", y = "Success (out of 100)", fill = "Success strategy") +
  theme_minimal()
```

## Plotting means

```{r}
note_mean  <- ggplot2::ggplot(notebook_tib, aes(film, arousal))
note_mean +
  stat_summary(fun.y = "mean", geom = "point", size = 3)+
  coord_cartesian(ylim = c(0, 50)) +
  scale_y_continuous(breaks = seq(0, 50, 5)) +
  labs(x = "Type of film", y = "Arousal (out of 100)") +
  theme_minimal()
```

`
```{r}
note_mean +
  stat_summary(fun.data = "mean_cl_normal", geom = "pointrange") +
  coord_cartesian(ylim = c(0, 50)) +
  scale_y_continuous(breaks = seq(0, 50, 5)) +
  labs(x = "Type of film", y = "Arousal (out of 100)") +
  theme_minimal()
```

```{r}
ggplot2::ggplot(notebook_tib, aes(film, arousal)) +
  stat_summary(fun.data = "mean_cl_normal", geom = "pointrange") +
  coord_cartesian(ylim = c(0, 50)) +
  scale_y_continuous(breaks = seq(0, 50, 5)) +
  labs(x = "Type of film", y = "Arousal (out of 100)") +
  facet_wrap(~sex) +
  theme_minimal()
```

```{r}
ggplot2::ggplot(notebook_tib, aes(film, arousal, colour = sex)) +
  stat_summary(fun.data = "mean_cl_normal", geom = "pointrange", position = position_dodge(width = 0.5)) +
  coord_cartesian(ylim = c(0, 50)) +
  scale_y_continuous(breaks = seq(0, 50, 5)) +
  labs(x = "Type of film", y = "Arousal (out of 100)", colour = "Biological sex") +
  theme_minimal()
```

## Violin plots

```{r}
ggplot2::ggplot(notebook_tib, aes(film, arousal, fill = sex)) +
  geom_violin() +
  stat_summary(fun.data = "mean_cl_normal", geom = "pointrange", position = position_dodge(width = 0.9)) +
  labs(x = "Biological sex", y = "Arousal", fill = "Film watched") +
  coord_cartesian(ylim = c(0, 50)) +
  scale_y_continuous(breaks = seq(0, 50, 5)) +
  theme_minimal()
```

```{r}
ggplot2::ggplot(notebook_tib, aes(film, arousal, fill = sex)) +
  geom_violin() +
  geom_boxplot(width = 0.2, position = position_dodge(width = 0.9)) +
  labs(x = "Biological sex", y = "Arousal", fill = "Film watched") +
  coord_cartesian(ylim = c(0, 50)) +
  scale_y_continuous(breaks = seq(0, 50, 5)) +
  theme_minimal()
```

## Repeated measures

```{r}
hiccups_tib %>% 
  dplyr::arrange(id)
```

```{r}
ggplot2::ggplot(hiccups_tib, aes(intervention, hiccups)) +
  stat_summary(fun.data = "mean_cl_normal", geom = "pointrange") +
  labs(x = "Intervention", y = "Number of hiccups") +
  coord_cartesian(ylim = c(0, 20)) +
  scale_y_continuous(breaks = seq(0, 20, 5)) +
  theme_minimal()
```

## Line plots and mixed designs

```{r}
grammar_tib %>% 
  dplyr::arrange(id)
```

```{r}
ggplot2::ggplot(grammar_tib, aes(time, grammar, colour = media_use)) +
  stat_summary(fun.data = "mean_cl_normal", geom = "pointrange", position = position_dodge(width = 0.05)) +
  stat_summary(fun.y = "mean", geom = "line", position = position_dodge(width = 0.05), aes(group = media_use)) +
  labs(x = "Time", y = "Grammar score (%)", colour = "Media use") +
  coord_cartesian(ylim = c(0, 75)) +
  scale_y_continuous(breaks = seq(0, 75, 5)) +
  theme_minimal()
```


## Scatterplots

### Simple scatterplot

```{r}
ggplot(exam_tib, aes(anxiety, exam_grade)) +
  geom_point() +
  labs(x = "Exam anxiety", y = "Exam performance (%)") +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
  scale_x_continuous(breaks = seq(0, 100, 10)) +
  theme_minimal()
```

### Catterplots

```{r}
ggplot(cat_tib, aes(dinner_time, meow)) +
  geom_point(colour = "#d35400", size = 2) +
  scale_y_continuous(breaks = seq(0, 16, 2)) +
  scale_x_continuous(breaks = seq(0, 24, 2)) +
  labs(x = "Time since dinner (hours)", y = "Purr volume") +
  theme_minimal()
```

### Grouped scatteplots

```{r}
ggplot(exam_tib, aes(anxiety, exam_grade, colour = sex))  +
  geom_point() +
  geom_smooth(method = "lm", aes(fill = sex), alpha = 0.1) +
  labs(x = "Exam anxiety", y = "Exam performance (%)", colour = "Biological sex", fill = "Biological sex") +
  coord_cartesian(ylim = c(0, 130)) +
  scale_y_continuous(breaks = seq(0, 130, 10)) +
  scale_x_continuous(breaks = seq(0, 100, 10)) +
  theme_minimal()
```

### Matrix scatterplots

```{r}
GGally::ggpairs(ong_tib, columns = c("rating", "extraversion", "narcissism"))
```

```{r}
GGally::ggpairs(ong_tib, columns = c("rating", "extraversion", "narcissism"), mapping = aes(colour = rating_type, alpha = 0.1)) + theme_minimal()
```


