---
title: "data"
author: "Andy Field"
output: learnr::tutorial
---

To work *outside of this tutorial* you need to download the following data file:

* [dog_training.csv](https://www.discovr.rocks/csv/dog_training.csv)
* [exercise.csv](https://www.discovr.rocks/csv/exercise.csv)

Set up an `r rstudio()` project in the way that [I recommend in this video](https://youtu.be/mqT7c17tofE?si=R0aSApz66fGQKESl), and save the data files to the folder within your project called [data]{.alt}. Place this code in the first code chunk in your `r quarto()` document:

```{r, eval=FALSE}
train_tib <- here::here("data/dog_training.csv") |> 
  read_csv() |> 
```

For the exercise data, import using this code:

```{r, eval=FALSE}
exercise_tib <-  here::here("data/exercise.csv") |> 
  read_csv() |>
  dplyr::mutate(
    id = as_factor(id),
    intervention = as_factor(intervention) |>  fct_relevel("Wait list"),
    time = as_factor(time) |> fct_relevel("Baseline", "1 month", "6 months", "12 months")
  )
```

This code reads in the data and converts the variables `intervention` and `time` to factors (categorical variable). It uses `fct_relevel` to set the order of the levels of the factor `intervention` to be wait list and intervention. Similarly, the order of the levels of the factor `time` are set as baseline, 1 month, 6 months, 12 months.