---
title: "data"
author: "Andy Field"
output: learnr::tutorial
---

To work *outside of this tutorial* you need to download the following data file:

* [goggles.csv](https://www.discovr.rocks/csv/goggles.csv)
* [xbox.csv](https://www.discovr.rocks/csv/xbox.csv)

Set up an `r rstudio()` project in the way that [I recommend in this video](https://youtu.be/mqT7c17tofE?si=R0aSApz66fGQKESl), and save the data files to the folder within your project called [data]{.alt}. Place this code in the first code chunk in your `r quarto()` document:

```{r, eval=FALSE}
goggles_tib <- here::here("data/goggles.csv") |>
  readr::read_csv() |>
  dplyr::mutate(
    facetype = as_factor(facetype),
    alcohol = as_factor(alcohol)
  )
```

This code reads in the data and converts the variables `alcohol` and `facetype` to factors (categorical variable). 

For the xbox data use this code (in which I have been explicit in setting certain categories as the first level of the factor using `fct_relevel()`):

```{r, eval=FALSE}
xbox_tib <- here::here("data/xbox.csv") |>
  read_csv() |>
  mutate(
    game = as_factor(game) |> fct_relevel("Static"),
    console = as_factor(console) |> fct_relevel("Xbox One")
  )
```