---
title: "packages"
author: "Andy Field"
output: learnr::tutorial
---


### Packages

```{r, eval = F, echo = F}
# Create bib file for R packages

tidylist <- c('tidyverse', 'dplyr', 'forcats', 'ggplot2', 'readr', 'stringr', 'tibble', 'tidyr')
easylist <- c("easystats", "datawizard", "insight")
otherlist <- c('here', 'qqplotr')

here::here("inst/tutorials/discovr_06/packages.bib") |>
  knitr::write_bib(c(easylist, tidylist, otherlist), file = _)
```

In this tutorial all packages are pre-loaded. If working outside of the tutorial, load the `easystats` and `tidyverse` packages (and any others if you're using a concise style - see previous section) at the beginning of your `r quarto()` document:

```{r eval = FALSE}
library(easystats)
library(tidyverse)
```

This tutorial uses the following packages:

- `here` [@R-here]
- `qqplotr` [@qqplotr2018; @R-qqplotr]

It also uses these packages from:

### easystats

- `datawizard` [@R-datawizard; @datawizard2022
- `insight` [@R-insight]
 
### tidyverse 
 
- `dplyr` [@R-dplyr]
- `forcats` [@R-forcats]
- `ggplot2` [@R-ggplot2; @wickham_ggplot2_2016]
- `readr` [@R-readr]
- `stringr` [@R-stringr]
- `tibble` [@R-tibble]
- `tidyr` [@R-tidyr]
