---
title: "data"
author: "Andy Field"
output: learnr::tutorial
---

To work *outside of this tutorial* you need to download the following data file:

* [biggest_liar.csv](https://www.discovr.rocks/csv/biggest_liar.csv)
* [exam_anxiety.csv](https://www.discovr.rocks/csv/exam_anxiety.csv)

Set up an `r rstudio()` project in the way that [I recommend in this video](https://youtu.be/mqT7c17tofE?si=R0aSApz66fGQKESl), and save the data files to the folder within your project called [data]{.alt}. Place this code in the first code chunk in your `r quarto()` document:

```{r, eval=FALSE}
liar_tib <- here::here("data/biggest_liar.csv") |> read_csv()
exam_tib <- here::here("data/exam_anxiety.csv") |> read_csv()
```

### Preparing data

To work *outside of this tutorial* you need to turn categorical variables into factors and set an appropriate baseline category using `as_factor` and `fct_relevel` from the `forcats` package. 

For [liar_tib]{.alt} execute the following code:

```{r, eval=FALSE}
liar_tib <- liar_tib |> 
  mutate(
    novice = as_factor(novice)
  )
```

For [exam_tib]{.alt} execute:

```{r, eval=FALSE}
exam_tib <- exam_tib |>
  mutate(
    id = as_factor(id),
    sex = as_factor(sex)
  )
```


