---
title: "packages"
author: "Andy Field"
output: learnr::tutorial
---


```{r, eval = F, echo = F}
# Create bib file for R packages

tidylist <- c('tidyverse', 'dplyr', 'ggplot2', 'readr', 'tibble')
easylist <- c("easystats", 'bayestestR', "datawizard", "insight", "parameters")
otherlist <- c('here', 'rstanarm')

here::here("inst/tutorials/discovr_bayes/packages.bib") |>
  knitr::write_bib(c(easylist, tidylist, otherlist), file = _)
```

In this tutorial all packages are pre-loaded. If working outside of the tutorial, load the `easystats` and `tidyverse` packages (and any others if you're using a concise style - see previous section) at the beginning of your `r quarto()` document:

```{r eval = FALSE}
library(easystats)
library(tidyverse)
```

This tutorial uses the following packages:

- `here` [@R-here]
- `rstanarm` [@R-rstanarm; @rstanarm2018]

It also uses these packages from:

### easystats

- `bayesR` [@R-bayestestR; bayestestR2019]
- `datawizard` [@R-datawizard; @datawizard2022]
- `insight` [@R-insight]
- `parameters` [@R-parameters; @parameters2020]
 
### tidyverse 
 
- `dplyr` [@R-dplyr]
- `ggplot2` [@R-ggplot2; @wickham_ggplot2_2016]
- `readr` [@R-readr]
- `tibble` [@R-tibble]
