---
title: "packages"
author: "Andy Field"
output: learnr::tutorial
---


```{r, eval = F, echo = F}
# Create bib file for R packages

tidylist <- c('tidyverse', 'dplyr', 'forcats', 'readr', 'tibble', 'tidyr')
easylist <- c('easystats', 'datawizard', 'effectsize', 'insight', 'parameters')
otherlist <- c('here', 'knitr', 'kableExtra')

here::here("inst/tutorials/discovr_19/packages.bib") |>
  knitr::write_bib(c(easylist, tidylist, otherlist), file = _)
```

In this tutorial all packages are pre-loaded. If working outside of the tutorial, load the [easystats]{.pkg} and [tidyverse]{.pkg} packages (and any others if you're using a concise style - see previous section) at the beginning of your `r quarto()` document:

```{r eval = FALSE}
library(easystats)
library(tidyverse)
```

This tutorial uses the following packages:

- [here]{.pkg} [@R-here]

It also uses these packages from:

### easystats 

- [datawizard]{.pkg} [@R-datawizard; @datawizard2022]
- [effectsize]{.pkg} [@R-effectsize; @effectsize2020]
- [insight]{.pkg} [@R-insight]
- [parameters]{.pkg} [@R-parameters; @parameters2020]
 
### tidyverse 
 
- [dplyr]{.pkg} [@R-dplyr]
- [forcats]{.pkg} [@R-forcats]
- [readr]{.pkg} [@R-readr]
- [tibble]{.pkg} [@R-tibble]
- [tidyr]{.pkg} [@R-tidyr].
