---
title: "Chapter 14"
author: "Andy Field"
date: '`r format(Sys.Date(), "%d %B %Y")`'
format:
  html:
    max-width: 3000px
    toc: true
execute:
  warning: false
  message: false
editor: visual
---

```{r}
library(broom)
library(broom.mixed)
library(magrittr)
library(dplyr)
library(ggplot2)
library(lme4)
library(lmerTest)

```

::: callout-important
## Usage

This file contains code relevant to chapter 14 of [Discovering Statistics Using R and RStudio](www.discovr.rocks) by [Andy Field](https://www.discoveringstatistics.com/about). These files contain abridged sections from the book so there are some copyright considerations but I offer them under a [Creative Commons Attribution-NonCommercial-NoDerivatives 4.0 International License](http://creativecommons.org/licenses/by-nc-nd/4.0/).[^1]
:::

[^1]: Basically you can use this material for teaching and non-profit activities but do not meddle with it or claim it as your own work.

## Load packages

Remember to load these packages:

```{r}
#| eval: false

library(dplyr)
library(ggplot2)

library(broom)
library(broom.mixed)
library(datawizard)
library(lme4)
library(lmerTest)

```

## Load the data

Load the data from the `discover` package:

```{r}
cosmetic_tib <- discovr::cosmetic
```

```{r}
#| eval: false
 
cosmetic_tib <- here::here("data-raw/csv_files/cosmetic.csv") |>

readr::read_csv() |>
  dplyr::mutate(
    clinic = forcats::as_factor(clinic),
    reason = forcats::as_factor(reason)
    )
```

If you want to read the file from the CSV and you have set up your project folder as I suggest in Chapter 1, then the code you would use is:

```{r}
#| eval: false
 
cosmetic_tib <- here::here("data/cosmetic.csv") |>
  readr::read_csv() |>
  dplyr::mutate(
    clinic = forcats::as_factor(clinic),
    reason = forcats::as_factor(reason)
    )
```

This code reads the file in and converts the variables **clinic** and **reason** to factors. It's a good idea to check that the levels of **clinic** are sequential from clinic 1 to clinic 21, and that **reason** has levels ordered change appearance and physical reason. Check the factor levels by executing:

```{r}

levels(cosmetic_tib$clinic)
levels(cosmetic_tib$reason)
```

If they're not in the correct order then:

```{r}
#| eval: false
 
cosmetic_tib <- cosmetic_tib |>
  dplyr::mutate(
    clinic = forcats::fct_relevel(clinic, paste("Clinic", seq(1, 20, 1))),
    reason = forcats::fct_relevel(reason, "Change appearance")
    )
```

## Plot the data

Create the plot

```{r}

ggplot2::ggplot(cosmetic_tib, aes(days, post_qol, colour = clinic)) +
  geom_point(alpha = 0.5, size = 1) +
  geom_smooth(method = "lm", se = FALSE, size = 0.5) +
  coord_cartesian(xlim = c(0, 400), ylim = c(0, 100)) +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
  labs(x = "Days post surgery", y = "Quality of life after surgery (%)", colour = "Clinic") +
  guides(colour=guide_legend(ncol=2)) +
  theme_minimal()
```

```{r}

ggplot2::ggplot(cosmetic_tib, aes(days, post_qol)) +
  geom_point(position = position_jitter(width = 0.1, height = NULL), alpha = 0.5, size = 1) +
  geom_smooth(method = "lm", se = FALSE, size = 0.5) +
  coord_cartesian(xlim = c(0, 400), ylim = c(0, 100)) +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
  labs(x = "Days post surgery", y = "Quality of life after surgery (%)", colour = "Clinic") +
  facet_wrap(~ clinic, ncol = 4) +
  theme_minimal()
```

## Summary statistics

```{r}

qol_sum <- cosmetic_tib |>
  dplyr::group_by(clinic) |>
  datawizard::describe_distribution(select = c("base_qol", "post_qol"), ci = 0.95)

qol_sum |>
  dplyr::mutate(
    Clinic = stringr::str_remove(.group, "clinic=")
    ) |>
  dplyr::select(Clinic, Variable:n) |>
  knitr::kable(caption = "Summary statistics for the surgery data",
               digits = 2)
```

## Fixed effect fits

### Fit the model to the pooled data

```{r}

pooled_lm <- lm(post_qol ~ days*reason + base_qol, data = cosmetic_tib)

broom::tidy(pooled_lm, conf.int = TRUE) |>
  knitr::kable(caption = "Parameter estimates for the pooled data model",
               digits = 3)
```

### Fit the model across contexts

```{r}

clinic_lms <- cosmetic_tib |>
  dplyr::arrange(clinic) |>
  dplyr::group_by(clinic) |>
  tidyr::nest() |>
  dplyr::mutate(
    model = purrr::map(.x = data,
                       .f = \(clinic_tib) lm(post_qol ~ days*reason + base_qol, data = clinic_tib)),
    coefs = purrr::map(model, tidy, conf.int = TRUE)
    )

models <- clinic_lms |>
  dplyr::select(-c(data, model)) |>
  tidyr::unnest(coefs)
```

```{r}

models |>
  ggplot(aes(estimate)) +
  geom_density() +
  facet_wrap(~term , scales = "free") +
  theme_minimal()
```

## Random effect models

### Try to fit the model

```{r}

cosmetic_mod <- lmerTest::lmer(post_qol ~ days*reason + base_qol + (days|clinic), data = cosmetic_tib)

allFit(cosmetic_mod)
```

### Rescale days

```{r}

cosmetic_tib <- cosmetic_tib |>
  dplyr::mutate(
    months = days*12/365
    )
```

### Standardizing variables

We don't use this model, but if you wanted to fit the model to standardized variables then use this code.

```{r}
#| eval: false


cosmetic_tib <- cosmetic_tib |>
  dplyr::mutate(across(.cols = where(is.double),
                       .fns = \(column) (column-mean(column, na.rm = T))/sd(column, na.rm = T),
                       .names = "z_{.col}")
                )

cosmetic_modz <- lmerTest::lmer(z_post_qol ~ z_days*reason + z_base_qol + (z_days|clinic), data = cosmetic_tib)

broom.mixed::tidy(cosmetic_modz)
```

### Fit the model using months

```{r}

cosmetic_mod <- lmerTest::lmer(post_qol ~ months*reason + base_qol + (months|clinic), data = cosmetic_tib, REML = T)

allFit(cosmetic_mod)
```

### Change the optimizer

```{r}

cosmetic_bob <- lmerTest::lmer(
  post_qol ~ months*reason + base_qol + (months|clinic),
  data = cosmetic_tib,
  control = lmerControl(optimizer="bobyqa")
  )
```

### Use maximum likelihood estimation

```{r}

cosmetic_ml <- lmerTest::lmer(
  post_qol ~ months*reason + base_qol + (months|clinic),
  data = cosmetic_tib,
  REML = F
  )
```

### Interpreting the model

```{r}
anova(cosmetic_bob) |> 
  knitr::kable(caption = "Fixed effects for the final model",
               digits = c(rep(2, 5), 3))
```

```{r}
options(knitr.kable.NA = '')
broom.mixed::tidy(cosmetic_bob, conf.int = T) |> 
  knitr::kable(caption = "Parameter estimates for the final model",
               digits = c(rep(2, 7), 3, 2, 2))
```

```{r}
cosmetic_slopes <- modelbased::estimate_slopes(cosmetic_bob,
                                               trend = "months",
                                               at = "reason",
                                               ci = 0.95) 

cosmetic_slopes |> 
  knitr::kable(caption = "Change over time for different reasons for surgery",
        digits = 3)
```

Some alternative ways to get the same simple slopes

```{r}

#| eval: false


cosmetic_slopes <- interactions::sim_slopes(
  cosmetic_bob,
  pred = months,
  modx = reason,
  confint = TRUE
  ) 

cosmetic_slopes <-  emmeans::emtrends(cosmetic_bob, var = "months", spec = "reason")
```

Plot the simple slopes

```{r}

interactions::interact_plot(
  model = cosmetic_bob,
  pred = months,
  modx = reason,
  interval = TRUE,
  x.label = "Months since surgery",
  y.label = "Quality of life post-surgery (0-100)",
  legend.main = "Reason for surgery"
  ) 
```
